### 2014-06-03 writing/S'14 paper/discussion.tex
Updating paper.

\subsection{General discussion}  % TODO

Discussion to be filled in based on Monday's results.\\...\\...\\...\\...\\...\\...\\...\\...\\...\\...

>>>>

\section{General discussion}

These experiments represent one of the first attempts to reproduce any large fragment of the behavior of a complex logic within a neural network model, and the first attempt that we are aware of to address either the encoding of lexical relations or the learning of recursive operators. This presents considerable challenges in evaluating the particular models that we choose, since we cannot rely on prior results to establish that any particular amount or type of training data is sufficient to teach any model the structure of the logic. The positive results that we have fonud, however, are extremely promising for the future of learned representation models in the applied modeling of meaning. We have seen that recursive neural tensor networks are able to encode lexical relations accurately and encode recursive operators. We have also seen that both RNNs and RNTNs are able to handle the meanings of quantifiers in an inference setting in at least some cases.

There is ample room to build on these results. In the interest of fully mirroring the capacity of existing natural logics in learned models, it would be valuable to extend these experiments to cover other ways in which meanings are encoded in natural language, including challenges such as reasoning over sentences with transitive verbs or relative clauses. In addition, it would be highly informative to compare these results on standard recursive neural networks with other proposed learned models for sentence meaning, such as dependency tree RNNs \cite{sochergrounded}, Belief Propagation RNNs (TODO: cite), or convolutional RNNs \cite{kalchbrenner2014convolutional}.

### 2014-06-03 writing/S'14 paper/intro.tex
Incremental paper update.
Adding Monday morning results to paper.
Updating paper.

Supervised recursive neural network models (RNNs) for sentence meaning
have been successful in a wide array of sophisticated language tasks,
including sentiment analysis, analogy completion, image description,
and paraphase detection. These results are encouraging about the
ability of these models to learn compositional semantic grammars, but
it remains an open question whether they can achieve the same results
as grammars based in logical forms when it comes to core semantic
concepts like quantification, entailment, and contradiction defined
over complex linguistic structures. These concepts are central to
robust natural language understanding, so it is essential that our
computational models be able to capture them.

We address this question directly by using a logical grammar to
generate controlled data sets encoding the semantic relationships
between pairs of expressions and evaluating whether each of two
classes of neural model --- plain RNNs and recursive neural tensor
networks (RNTNs) --- can learn those relationships correctly. Our
logical grammar is a version of the natural logic developed by
\cite{maccartney2009extended}, which defines seven core relations of
synonymy, entailment, contradiction, and mutual consistency. Natural
logic is well-suited to our purposes because of it defines a rich set
of relations between words, phrases, and sentences, and because its
logical expressions are derived directly from surface forms, making it
amenable to large-scale natural language processing.

In our first experiment, we show that both of these models can learn
the core relational algebra of natural logic from data. Our second
experiment builds on this result to cover relations between complex
recursive structures like `(A or B)' and `not(not A and not B)', and
our third experiment involves relations between quantified statements
like `every reptile walks' and `every turtle moves'. We find that the
plain RNN achieves only mixed results on the second and third
experiments, but that the stronger RNTN models generalized well in
every case, suggesting that it has in fact learned, or at least
learned to simulate, our target logical concepts.

\section{Introduction}
% TODO: cite \cite{watanabe2012latent} for past work on ML for McC&M-style NLI

@@

% TODO: Name RNTNs

@@

% TODO: Cite ICLR paper, emphasize new work since

\subsection{The task: natural language inference}

% Condense?
In the standard formulation of the inference task (and the one used in the RTE datasets \cite{dagan2006pascal}), the goal is to determine whether a reasonable human would infer a hypothesis from a premise.
MacCartney formalizes a method of inferring entailment relations, and moves past two way entailment/non-entailment classification, proposing the set $\mathfrak{B}$ of seven labels meant to describe all of the possible non-trivial relations that might hold between pairs of statements, shown in Table \ref{b-table}.

@@

In order to minimize this possibility, I define the task of strict unambiguous NLI (SU-NLI). In this task, only entailments that are licensed by a strictly literal interpretation of the provided sentences are considered valid, and several constraints are applied to the language to minimize ambiguity:
\begin{itemize}
\ex A small, unambiguous vocabulary is used.
\ex All strings are given an explicit unlabeled tree structure parse.
\ex Statements involving the hard-to-formalize generic senses of nouns---i.e. \ii{dogs bark} as opposed to the non-generic \ii{all dogs bark}---are excluded.
\ex The sentences do not contain context dependent elements. This includes any reference to time or any tense morphology, and all pronouns.
\ex The morphology is dramatically simplified: the copula is not used (\ii{some puppy is French} is simplified to \ii{some puppy French}, to make it more directly comparable to sentences like \ii{some puppy bark}), and agreement marking (\ii{they walk} vs. \ii{she walks}) is omitted.
\end{itemize}

The key to success at this task is to learn a set of representations and functions that can mimic the logical structure underlying the data. There is limited precedent that deterministic logical behavior can be learned in supervised deep learning models: \citet{socher2012semantic} show in an aside that a Boolean logic with negation and conjunction can be learned in a minimal recursive neural network model with one-dimensional (scalar) representations for words. Modeling the logical behavior that underlies linguistic reasoning, though, is a substantially more difficult challenge, even in modular hand-built models.

The natural logic engine at the core of MacCartney's \cite{maccartney2009natural} NLI system requires a complex set of linguistic knowledge, much of which takes the form of what he calls projectivity signatures. These signatures are tables showing the entailment relation that must hold between two strings that differ in a given way (such as the substitution of the argument of some quantifier), and are explicitly provided to the model
for dozens of different cases of insertion, deletion and substitution of different types of lexical item. For example in judging the inference \ii{no animals bark $|$ some dogs bark} it would first used stored knowledge to compute the relations introduced by each of the two differences between the sentences. Here, the substitution of \ii{no} for \ii{some}  yields $\natneg$ and the substitution of \ii{dogs} for \ii{animals} yields $\sqsupset$. It would then use an additional store of knowledge about relations to resolve the resulting series of relations into one ($|$) that expresses the relation between the two sentences being compared:
\begin{quote}

1. \ii{no animals bark $\natneg$ \textbf{some} animals bark}\\
2. \ii{some animals bark $\sqsupset$ some \textbf{dogs} bark}\\
3. \ii{no animals bark $[\natneg\bowtie\thinspace\sqsupset\thinspace = |]$ some dogs bark}

\end{quote}

This study is the first that I am aware of to attempt to build an inference engine based on learned vector representations, but two recent projects have attempted to introduce vector representations into inference systems in other ways: 
\citet{baroni2012entailment} have achieved limited success in building a classifier to judge entailments between one- and two-word phrases (including some with quantifiers), though their vector representations were crucially based on distributional statistics and were not  learned for the task.
In another line of research, \citet{garrette2013formal} propose a way to improve standard discrete NLI with vector representations. They propose a deterministic inference engine (similar to MacCartney's) which is augmented by a probabilistic component that evaluates individual lexical substitution steps in the derivation using vector representations, though again these representations are not learned, and no evaluations of this system have been published to date.
\label{sec2}

>>>>

Supervised recursive neural network models (RNNs) for sentence meaning
have been successful in a wide array of sophisticated language tasks,
including sentiment analysis \cite{socher2011semi}, analogy completion, image description,
and paraphase detection . These results are encouraging about the
ability of these models to learn compositional semantic grammars, but
it remains an open question whether they can achieve the same results
as grammars based in logical forms when it comes to core semantic
concepts like quantification, entailment, and contradiction defined
over complex linguistic structures. These concepts are central to
robust natural language understanding, so it is essential that our
computational models be able to capture them. % TODO: Insert more citations for past RNN work.

We address this question directly by using a logical grammar togenerate controlled data sets encoding the semantic relationships between pairs of expressions and evaluating whether each of twoclasses of neural model --- plain RNNs and recursive neural tensor networks (RNTNs, \cite{socher2013acl1}) --- can learn those relationships correctly. Following \citet{bowman2013can}, we use the natural logic developed by\cite{maccartney2009extended} as our formal model, which defines seven core relations ofsynonymy, entailment, contradiction, and mutual consistency. Natural logic is well-suited to our purposes because of it defines a rich setof relations between words, phrases, and sentences, and because its logical expressions are derived directly from surface forms, making itamenable to large-scale natural language processing. % Is this the right place for the first self citation?

In our first experiment, we show that both of these models can learn
the core relational algebra of natural logic from data. Our second
experiment builds on this result to cover relations between complex
recursive structures like `(A or B)' and `not(not A and not B)', and
our third experiment involves relations between quantified statements
like `every reptile walks' and `every turtle moves'. We find that the
plain RNN achieves only mixed results on the second and third
experiments, but that the stronger RNTN models generalized well in
every case, suggesting that it has in fact learned, or at least
learned to simulate, our target logical concepts. % TODO: Update with current results

Citations to additional past work to be added.\\...\\...\\...\\...\\...

% TODO: Mention earlier work on NN interpretation - \cite{Garcez-etal:2001}

@@

@@

\subsection{Natural language inference and natural logic}

In its simplest form, \ii{natural language inference} (also known as \ii{recognizing textual entailment}, as in \cite{dagan2006pascal}) is the task of determine whether one sentence entails another. Much of the theoretical work on this task (and some successful implemented models \cite{maccartney2009natural, watanabe2012latent}) involve \ii{natural logic}, formal systems that define sound rules of inference from one sentence of natural language to another without the need for intermediate representations in some other logic. The most powerful model that we are aware of for natural logic is due to MacCartney and Manning \cite{maccartney2009extended} and Icard \cite{icard2012inclusion}, and is centered around the definition of a set of seven logical relations which can hold between sentences, shown in Table \ref{b-table}.

This approach to natural logic can capture much of the complexity of natural language meaning within a well understood framework, and is also fairly straightforward to implement in a machine learning setting since it can be reduced to a seven-way classification problem on sentence pairs. Our goal in this paper is to learn recursive neural network models which are able to mimic key behaviors of this system.

@@

The goal of each of the three experiments that we propose is to learn classifiers that are able to classify pairs of statements from some highly constrained language into these seven classes. It should be noted that this is not a balanced classification problem. If arbitrary pairs of statements are chosen for comparison in almost any domain of natural langauge, the minimally informative \# relation will apply between them. % TODO: How much should we say about this?

% The natural logic engine at the core of MacCartney and Manning's system requires a complex set of linguistic knowledge, much of which takes the form of what he calls projectivity signatures. These signatures are tables showing the entailment relation that must hold between two strings that differ in a given way (such as the substitution of the argument of some quantifier), and are explicitly provided to the model
%for dozens of different cases of insertion, deletion and substitution of different types of lexical item. For example in judging the inference \ii{no animals bark $|$ some dogs bark} it would first used stored knowledge to compute the relations introduced by each of the two differences between the sentences. Here, the substitution of \ii{no} for \ii{some}  yields $\natneg$ and the substitution of \ii{dogs} for \ii{animals} yields $\sqsupset$. It would then use an additional store of knowledge about relations to resolve the resulting series of relations into one ($|$) that expresses the relation between the two sentences being compared:
%\begin{quote}

%1. \ii{no animals bark $\natneg$ \textbf{some} animals bark}\\
%2. \ii{some animals bark $\sqsupset$ some \textbf{dogs} bark}\\
%3. \ii{no animals bark $[\natneg\bowtie\thinspace\sqsupset\thinspace = |]$ some dogs bark}

%\end{quote}

Work to date on inference in neural network models is quite limited.
\citet{baroni2012entailment} have achieved limited success in building a classifier to judge entailments between one- and two-word phrases (including some with quantifiers), though their vector representations were crucially based on distributional statistics and were not  learned for the task.
In another line of research, \citet{garrette2013formal} propose a way to improve standard discrete NLI with vector representations. They propose a deterministic inference engine (similar to MacCartney's) which is augmented by a probabilistic component that evaluates individual lexical substitution steps in the derivation using vector representations, though again these representations are not learned, and no evaluations of this system have been published to date.
\label{sec2}

### 2014-06-03 writing/S'14 paper/join.tex
Updating paper.

I test the model's ability to learn this behavior by creating an artificial dataset of terms which represent sets of numbers. Since MacCartney and Manning's set of relations hold between sets as well as between sentences, I can use the underlying set structure to generate the all of the relations that hold between any pair of these terms, as in Figure \ref{lattice-figure}. I train the model defined above on a subset of these relations, but rather then presenting the model with a pair of tree-structured sentences as inputs, simply present it with two single terms, each of which corresponds to a single vector in the (randomly initialized) vocabulary matrix $V$, ensuring that the model has no information about the terms being compared except the relations between them.

I generate 80 randomly generated sets drawing from the same seven elements, and create a dataset consisting of the relations between every pair of sets, yielding 6400 pairs. 3200 of these pairs were then chosen as a test dataset, and that test dataset was further split into the 2960 examples that can be provably derived from the test data using MacCartney and Manning's join table (or by the symmetry of the relations in about half of the cases) and the 240 that cannot. % TODO: Say more about symmetry?

We found that the RNTN model worked best with 11 dimensional vector representations for the 80 sets and a 90 dimensional feature vector for the classifier. This model was able to correctly label 99.3\% of the derivable test examples, and 99.1\% of the remaining examples. The simpler RNN model worked best with 11 and 75 dimensions, respectively, but was able to achieve accuracies of only 90.0\% and 87.\%, respectively.

Discussion to be filled in based on Monday's results.\\...\\...\\...\\...\\...\\...

% Points for discussion: 
% - Fairly simple interpretation: RNTN generalized, RNN didn't
% - Future work: Can this generalize to models with huge numbers of entities, or does this result depend on the examples reflecting a small model in any way? Can't exhaustively generate large models without getting \#s - need sampling.

>>>>

We test the model's ability to learn this behavior by creating an artificial dataset of terms which represent sets of numbers. Since MacCartney and Manning's set of relations hold between sets as well as between sentences, we can use the underlying set structure to generate the all of the relations that hold between any pair of these terms, as in Figure \ref{lattice-figure}. We train the model defined above on a subset of these relations, but rather then presenting the model with a pair of tree-structured sentences as inputs, simply present it with two single terms, each of which corresponds to a single vector in the (randomly initialized) vocabulary matrix $V$, ensuring that the model has no information about the terms being compared except the relations between them.

We generate 80 randomly generated sets drawing from the same seven elements, and create a dataset consisting of the relations between every pair of sets, yielding 6400 pairs. 3200 of these pairs were then chosen as a test dataset, and that test dataset was further split into the 2960 examples that can be provably derived from the test data using MacCartney and Manning's join table (or by the symmetry of the relations in about half of the cases) and the 240 that cannot. % TODO: Say more about symmetry?

We tested a version of both the RNN model and the RNTN model on these data. In both cases, the models were implemented exactly as described in Section \ref{methods}, but since the items being compared are single terms rather than full tree structures, the composition layer was not used, and the two models differed only in which layer function was used for the comparison layer.
We found that the RNTN model worked best with 11 dimensional vector representations for the 80 sets and a 90 dimensional feature vector for the classifier. This model was able to correctly label 99.3\% of the derivable test examples, and 99.1\% of the remaining examples. The simpler RNN model worked best with 11 and 75 dimensions, respectively, but was able to achieve accuracies of only 90.0\% and 87.\%, respectively.

% TODO: T-sne the RNTN model's embeddings.

These results are fairly simple to interpret. The RNTN model was able to accurately encode the relations between the terms in the geometric relations between their vectors, and was able to then use that information to recover relations that were not overtly included in the training data. The RNN model was able to approximate this behavior only incompletely. It is possible but not likely that it could be made to find a good solution with further optimization on different learning algorithms, or that it would do better on a larger universe of sets for which there would be a larger set of training data to learn from.

### 2014-06-03 writing/S'14 paper/methods.tex
Adding Monday morning results to paper.
Updating paper.

\section{Recursive neural network models}

The models that we study is are centered aronund a recursively applied composition function which is meant to mimic the recursive construction of meanings in formal semantics. In this scheme, pairs of words are merged into phrase representations by a function that maps from representations of length $2N$ to representations of length $N$. These phrase representations are then further merged with other words and phrases until the entire phrase or sentence being evaluated is represented in a single vector. This vector is then used as the input to a classifier and used in a supervised learning task.

Borrowing a model directly from the existing literature for this task is impossible since none has been proposed to detect asymmetric relations between phrases. Instead, we build a combination model, depicted in Figure \ref{sample-figure}. The two phrases being compared are built up separately on each side of the tree using the same composition function until they have each been reduced to single vectors. Then, the two phrase vectors are fed into a separate comparison layer that is meant to generate a feature vector capturing the relation between the two phrases. The output of this layer is then given to a softmax classifier, which in turn produces a hypothesized distribution over the seven relations.

For a composition layer, we experiment with both a plain neural network layer function (eq. \ref{rnn}) and the RNTN layer function proposed in \citet{chen2013learning} (eq. \ref{rntn}). A sigmoid nonlinearity (elementwise $\tanh$) is applied to the output of either layer function, following Socher et al.
\begin{gather} \label{rnn}
y_{RNN} = f(\mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + b_i)\\ % TODO: Add column vectors?
\label{rntn}
y_{RNTN} = f(\vec{x}^{(l)T} \mathbf{A}^{[1...N]} \vec{x}^{(r)} + \mathbf{B} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{c})
\end{gather} % TODO: Explain third order tensor parameter?

The comparison layer uses the same type of function with different parameters and a different nonlinearity function wrapped around it. Rather than use a $\tanh$ nonlinearity here, I found a substantial improvement in performance by using a rectified linear function for $f_{b}$. In particular, I use the leaky rectified linear function \cite{maasrectifier}: $f_{b}(\vec{x})=\max(\vec{x}, 0) + 0.01\min(\vec{x}, 0)$,  applied elementwise.

To run the model forward and label a pair of phrases, the structure of the lower layers of the network is assembled so as to mirror the tree structures provided for each phrase. The word vectors are then looked up from the vocabulary matrix $V$ (one of the model parameters), and the composition and comparison functions are used to pass information up the tree and into the classifier at the top. The model is trained using backpropagation through structure  \cite{goller1996learning}, wherein the negative log of the probability assigned to the correct label is taken as a cost function, and the gradient of each parameter with respect to that cost function is computed at each node, with information passing down the tree and into both the function parameters and the vocabulary matrix. Gradients for different instances of the composition function or different instances of a word in the same tree are summed to produce at most a single gradient update per parameter.

\textit{Optimization:} I train the model with stochastic gradient descent (SGD), with gradients pooled from randomly chosen minibatches of 32 training examples, and learning rates computed using AdaGrad \cite{duchi2011adaptive} from a starting rate of 0.2. The parameters (including the vocabulary) are initialized randomly using a uniform distribution over $[-0.1, 0.1]$. % TODO: Update

\begin{figure}
\begin{center}
% \Tree [.{\sc softmax classifier}  [.{\sc comparison layer} [.{\sc \ii{dog} vector from $V$} ] [.{\sc composition layer} [.{\sc composition layer} [.{\sc \ii{very} vector from $V$} ] [.{\sc \ii{big} vector from $V$} ] ] [.{\sc \ii{cat} vector from $V$} ] ] ] ]
\includegraphics[scale=0.36]{model.eps}
\end{center}
\caption{The model structure used to compare \ii{no dog} and \ii{(not all) dog}. \label{sample-figure}} 
\end{figure} % TODO: Create a more compact figure using something better than qtree.

\ii{Source code and data will be released after the conclusion of the review period.} % TODO: Or upon request? Attach anonymized code?

>>>>

\section{Recursive neural network models} \label{methods}

The models that we study are are centered aronund a recursively applied composition function which is meant to mimic the recursive construction of meanings in formal semantics. In this scheme, pairs of words are merged into phrase representations by a function that maps from representations of length $2N$ to representations of length $N$. These phrase representations are then further merged with other words and phrases until the entire phrase or sentence being evaluated is represented in a single vector. This vector is then used as the input to a classifier and used in a supervised learning task.

We use the model architecture proposed in \citet{bowman2013can} and depicted in Figure \ref{sample-figure}. The two phrases being compared are built up separately on each side of the tree using the same composition function until they have each been reduced to single vectors. Then, the two phrase vectors are fed into a separate comparison layer that is meant to generate a feature vector capturing the relation between the two phrases. The output of this layer is then given to a softmax classifier, which in turn produces a hypothesized distribution over the seven relations.

For a composition layer, we evaluate models with both the plain neural network layer function (eq. \ref{rnn}) and the RNTN layer function proposed in \citet{chen2013learning} (eq. \ref{rntn}). A sigmoid nonlinearity (elementwise $\tanh$) is applied to the output of either layer function, following Socher et al.
\begin{gather} \label{rnn}
y_{RNN} = f(\mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + b_i)\\ % TODO: Add column vectors?
\label{rntn}
y_{RNTN} = f(\vec{x}^{(l)T} \mathbf{A}^{[1...N]} \vec{x}^{(r)} + \mathbf{B} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{c})
\end{gather} % TODO: Explain third order tensor parameter?

The comparison layer uses the same type function template as the composition function (either an NN layer or an NTN layer) with independantly learned parameters and a separate nonlinearity function. Rather than use a $\tanh$ nonlinearity here, we found a substantial improvement in performance by using a rectified linear function for $f_{b}$. In particular, we use the leaky rectified linear function \cite{maasrectifier}: $f_{b}(\vec{x})=\max(\vec{x}, 0) + 0.01\min(\vec{x}, 0)$,  applied elementwise.

To run the model forward and label a pair of phrases, the structure of the lower layers of the network is assembled so as to mirror the tree structures provided for each phrase. The word vectors are then looked up from the vocabulary matrix $V$ (one of the model parameters), and the composition and comparison functions are used to pass information up the tree and into the classifier at the top. For an objective function, we use the negative log of the probability assigned to the correct label.

We train the model with minibatch stochastic gradient descent (SGD) with learning rates computed using AdaGrad \cite{duchi2011adaptive} from a starting rate of 0.2. The parameters (including the vocabulary) are initialized randomly using a uniform distribution over $[-0.1, 0.1]$. % TODO: Update

\begin{figure}
\begin{center}
\includegraphics[scale=0.35]{model.eps}
\end{center}
\caption{The model structure used to compare \ii{no dog} and \ii{(not all) dog}. The same structure is used for both the RNN and RNTN layer functions. \label{sample-figure}} 
\end{figure}

\ii{Source code and generated data will be released after the conclusion of the review period.} % TODO: Or upon request? Attach anonymized code?

### 2014-06-03 writing/S'14 paper/quantifiers.tex
Updating paper.

\caption{Quantifier experiment performance. *Most frequent class accuracy is measured against the most frequent class in the training data, \#.\label{resultstable}}
\end{table} % TODO: Replace

@@

Discussion to be filled in based on Monday's results.\\...\\...\\...\\...\\...\\...\\...\\...\\...\\...

% Notes for discussion:
% - First effort to learn a logic, results somewhat unclear.
% -- No straightforward way to prove that there is enough information in the data to learn quantifier projecitivies in this setting
% - Perfect performance on no-no promising that the model is learning at least some structure

% TODO: Revise

>>>>

\caption{Quantifier experiment performance. *Most frequent class accuracy is measured against the most frequent class in the training data, \#.\label{resultstable}}
\end{table}

@@

 (Note: Models still running. Results might improve and this para is likely to change.) The strong performance by the RNN on the \textsc{pair no/no} experiment suggests that it is at least plausible that this model is able to handle quantifiers correctly given sufficient training data, but the sub 90\% results on most of the training settings suggest that the data provided in these experiments are not generally sufficient, and that the model may not be able to learn to reason with quantifiers with high accuracy. However, the fact that both models perform far above baseline is promising, and the question of how much data is necessary to accurately capture quantifier behavior in a na\"ive model remains open.

Discussion of RNN results (and updated discussion of better RNTN results) to be filled in as results arrive.\\...\\... % TODO

### 2014-06-03 writing/S'14 paper/recursion.tex
Updating paper.

% NOTE: Worth explicitly calling this project theorem proving? Yes! There was some confusion at CSLI.

Socher et al. \cite{socher2012semantic} have previously demonstrated the learning of a logic in a matrix-vector RNN model somewhat similar to our own, but the logic discussed here is substantially stronger, and a much better approximation of the kind of structure that is needed for natural language. The logic learned in that experiment is boolean, wherein the atomic symbols are simply the values 0 and 1, rather than variables over those values. While learning the operators of that logic is not trivial, the ouptuts of each operator can be represented accurately by a single bit. The statements of propositional logic learned here describe conditions on the truth values of propositions where those truth values are not known. As opposed to the two-way contrasts seen in \cite{socher2012semantic}, this logic distinguishes between 64 (2^6) possible assignments of truth values, and expressions of this logic define arbitrary conditions on these possible assignments, for a total of 2^{64} ($\approx 10^{20}$) possible statements that the recursive model needs to be able to distinguish. To frame this distinction in another way, the relational statements in our data our theorems about the relations between statements in the logic tested in \cite{socher2012semantic}.

@@

Discussion to be filled in based on Monday's results.\\...\\...\\...\\...\\...\\...\\...\\...\\...\\...

% TODO:
% Points for discussion
% - The RNTN has learned to approximate the correct funtionn
% -- The error introduced by the approximation likely grows with depth, but is sufficiently small at the training example sizes

% - The RNN has plausibly learned an approximation as well, but one which is quite noisy even without recursion
% -- Possible but unlikely that further tuning and different types or amounts of training data could get the RNN to work

% - Future work:
% -- Generate much larger examples to tease apart the ability of the model to approximate a representable function from data, and the ability of the model to represent

>>>>

% NOTE: There was some confusion at CSLI and the 224U lecture about whether this was theorem proving or some kind of model checking. Hopefully this is explicit enough.
Socher et al. \cite{socher2012semantic} have previously demonstrated the learning of a logic in a matrix-vector RNN model somewhat similar to our own, but the logic discussed here is substantially stronger, and a much better approximation of the kind of structure that is needed for natural language. The logic learned in that experiment is boolean, wherein the atomic symbols are simply the values 0 and 1, rather than variables over those values. While learning the operators of that logic is not trivial, the ouptuts of each operator can be represented accurately by a single bit. The statements of propositional logic learned here describe conditions on the truth values of propositions where those truth values are not known. As opposed to the two-way contrasts seen in \cite{socher2012semantic}, this logic distinguishes between 64 (2^6) possible assignments of truth values, and expressions of this logic define arbitrary conditions on these possible assignments, for a total of 2^{64} ($\approx 10^{20}$) possible statements that the recursive model needs to be able to distinguish. To frame this distinction in another way, the relational statements in our data our theorems about the relations between statements in the logic tested in \cite{socher2012semantic}.

@@

The performance of the RNTN model on small unseen test examples indicates that it learned a correct approximation of the underlying logic. It appears that this approximation is accurate enough to yield correct answers when the composition layer is only applied a small number of times, but that the error in the approximation grows with depth, resulting in the gradually dropping performance. This is not necessarily a significant flaw in the model, since it remains possible that a less lossy approximation of the same logic would do no better on the training objective, since that objective only considers short strings and includes a regularization term. Whether a more accurate encoding could be learned from a larger set of training data remains a quention for future work. In contrast, the RNN model did not appear to be able to learn a correct approximation of the logic.
